Hypertension and microbiome

### 2020-03-12 articletwo.Rmd
Lactobacillus excretition instead of concentration

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(egg)
library(gridExtra)
library(png)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
library(pander)
library(hexbin)
```

@@

```{r Functions}
sourcefiles <- c("articletwo-officer.R",
                 "articletwo-rrbiome.R",
                 "articletwo-ggplot.R")
```

@@

 ```{r adonis calculation}
 if (!file.exists("rds/adonis.species.rds")) {
     adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                               matrix = bray.dist.m.species,
                                               vars = list("max" = var.CL,
                                                            "min" = var.CL.min))
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
 } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
 }
```

@@

```{r lacto deseq calculation}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("NA.", var.CL),
                    saltsubset = TRUE)
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
```

Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`.

```{r significant results for salt subset}
dds.salt %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH")) %>%
    filter(row == "Lactobacillus (Bacteria)") %>%
    kable
```

@@

```{r scatter plot from deseq, echo = FALSE}
g.salt <- saltboxplot(pseq.genus.salt, dds.salt)
ggsave(file = "cache/gsalt.png", plot = g.salt, height = 5, width = 5, dpi = 300)
```

<img src="cache/gsalt.png" />

@@

```{r species salt}
dds.salt.species <- myDESeq(pseq = pseq.species,
                        vars = c("NA.", var.CL),
                        coreterm = "^Lactobacillus.*Bacteria",
                        saltsubset = TRUE)
```

@@

```{r species salt results}
dds.salt.species %>%
    results(., name = "NA.", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
                  row = renametaxa(row)) %>%
    filter(qval < 0.05) %>%
    kable
```

@@

```{r write supplement}
core.names.incols <- pseq.genus.coretaxa %>%
    renametaxa %>%
    sort %>%
    list.partition %>%
    mycbind %>%
    as.data.frame

@@

```{r write supplement}
lactobacilli.names.incols <- pseq.species %>%
    taxa %>%
    mygrep(word = "^Lactobacillus.*Bacteria") %>%
    renametaxa %>%
    sort %>%
    list.partition(parts = 4) %>%
    mycbind %>%
    as.data.frame

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(tibble)
library(phyloseq)
library(nortest)
library(microbiome)
library(knitr)
library(tidyr)
library(vegan)
library(reshape)
library(parallel)
library(officer)
library(flextable)
library(xtable)
library(rvg)
library(tableone)
library(scales)
library(ggplot2)
library(egg)
library(gridExtra)
library(png)
library(ggpubr)
library(broom)
library(ggfortify)
library(RColorBrewer)
library(gvlma)
library(purrr)
library(gtable)
library(car)
library(M3C)
library(emmeans)
library(DESeq2)
library(ggeffects)
library(pander)
library(hexbin)
library(readr)
```

@@

```{r Functions}
sourcefiles <- c("articletwo-officer.R",
                 "articletwo-import.R",
                 "articletwo-rrbiome.R",
                 "articletwo-ggplot.R")
```

@@

 ```{r adonis calculation}
 if (!file.exists("rds/adonis.species.rds")) {
     adonis.species <- calculate.betadiversity(pseq = pseq.species,
                                               matrix = bray.dist.m.species,
                                               npermutations = 999,
                                               vars = list("max" = var.CL,
                                                            "min" = var.CL.min))
     saveRDS(adonis.species, file = "rds/adonis.species.rds")
 } else {
     adonis.species  <- readRDS("rds/adonis.species.rds")
 }
```

@@

```{r lacto deseq calculation}
pseq.genus.salt <- pseqsubset(pseq.genus, coretaxa = pseq.genus.coretaxa, saltsubset = TRUE)
```

Lactobacillus prevalence in subsample `r sprintf("%.1f%%", 100*pseq_prevalence(pseq.genus.salt))`. 
Levels of dU-Na `r sprintf("%.1f (%.1f)", mean(meta(pseq.genus)$dUNA, na.rm = TRUE), 
	sd(meta(pseq.genus)$dUNA, na.rm = TRUE))`

```{r significant results for salt subset}
dds.salt <- myDESeq(pseq = pseq.genus,
                    coretaxa = pseq.genus.coretaxa,
                    vars = c("dUNA", var.CL),
                    saltsubset = TRUE)

dds.salt %>%
    results(., name = "dUNA", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
                  lfc_se = sprintf("%.4f±%.4f", log2FoldChange, lfcSE)) %>%
    select(row, lfc_se, qval) %>%
    filter(row == "Lactobacillus (Bacteria)") %>%
    kable
```

@@

```{r salt subset char}
df.genus.salt <- saltdset(pseq.genus.salt, dds.salt)
```

```{r genus salt levels}
df.genus.salt %>%
    group_by(group) %>%
    summarize(mean = round(mean(dUNA), 1),
              sd = round(sd(dUNA), 1)) %>%
    kable
```

Mean urinary salt `r df.genus.salt$dUNA %>% mean` and sd `r df.genus.salt$dUNA %>% sd`.

@@

```{r species salt}
dds.salt.species <- myDESeq(pseq = pseq.species,
                        vars = c("dUNA", var.CL),
                        coreterm = "^Lactobacillus.*Bacteria",
                        saltsubset = TRUE)
```

@@

```{r species salt results}
dds.salt.species %>%
    results(., name = "dUNA", tidy = TRUE) %>%
    dplyr::mutate(qval = p.adjust(pvalue, method="BH"),
                  row = renametaxa(row)) %>%
    filter(qval < 0.05) %>%
    mutate(qval_exp = sprintf("%.4e", qval)) %>%
    kable
```

Prevalence for L. paracasei is
`r pseq_prevalence_format(pseq.species, "Lactobacillus_paracasei (Bacteria)")`.
and for L. salivarius is 
`r pseq_prevalence_format(pseq.species, "Lactobacillus_salivarius (Bacteria)")`.

```{r species plot}
df.species.salt <- deseq.abundances(dds.salt.species) %>%
    full_join(deseq.abundances(dds.salt), by = "sampleid") %>%
    left_join(meta(pseq.species) %>% tibble::rownames_to_column("sampleid"), by = "sampleid") %>%
    mutate_at(vars(contains('Lactobacillus')), .funs = list(group = ~factor(dplyr::ntile(., 4)))) %>%
    select(dUNA,
           Lactobacillus_paracasei.Bacteria_group,
           Lactobacillus_salivarius.Bacteria_group,
           Lactobacillus.Bacteria_group)

df.species.salt %>%
    saltboxplot %>%
    gather(taxa, abundance, -dUNA) %>%
    ggsave(file = "cache/boxplot-lacto-salt.png", plot = ., height = 3, width = 3.5, dpi = 300, unit = "in")
```

<img src = "cache/boxplot-lacto-salt.png" />

Urinary sodium quartiles for L. paracei

```{r species salt levels paracasei, echo = FALSE}
df.species.salt %>%
    group_by(Lactobacillus_paracasei.Bacteria_group) %>%
    summarize(mean = round(mean(dUNA), 1),
              sd = round(sd(dUNA), 1)) %>%
    kable
```

Urinary sodium quartiles for L. salivarius

```{r salt levels salivarius, echo = FALSE}
df.species.salt %>%
    group_by(Lactobacillus_salivarius.Bacteria_group) %>%
    summarize(mean = round(mean(dUNA), 1),
              sd = round(sd(dUNA), 1)) %>%
    kable
```

@@

```{r write supplement core names}
core.names.incols <- pseq.genus.coretaxa %>%
    renametaxa %>%
    sort %>%
    list.partition %>%
    mycbind %>%
    as.data.frame

@@

```{r write supplement lactobacilli}
lactobacilli.names.incols <- pseq.species %>%
    taxa %>%
    mygrep(word = "^Lactobacillus.*Bacteria") %>%
    renametaxa %>%
    sort %>%
    list.partition(parts = 4) %>%
    mycbind %>%
    as.data.frame
